---
title: "**ctmm analysis mammals**"
subtitle: "**Supplmentary Material**"
author: "Martinig, A. R., and M. J. Noonan"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "**Table of Contents**"
    output-file: "index.html"
    theme:
      dark: darkly
      light: flatly
    embed-resources: true
    code-fold: true
    code-tools: true
    number-sections: true
    #bibliography: ./bib/ref.bib
    fontsize: "12"
    max-width: "10"
    code-overflow: wrap
crossref: 
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Tab.    # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
  #cache: true
---

## Setting up

### Load packages

```{r load packages}
#force install pacman

#install.packages("pacman")
#pacman::p_load(devtools, tidyverse, metafor, patchwork, R.rsp, emmeans)

#load all other packages
pacman::p_load(tidyverse, # tidy family and related pacakges below
               kableExtra, # nice tables
               ctmm,  
               fitdistrplus,
               metRology,
               DAAG, 
               ggdist,
               lme4,
               mgcv,
               ggplot2, # post-hoc tests
               gridExtra, 
               pander,   # nice tables
               metafor,  # package for meta-analysis
               ape,      # phylogenetic analysis
               ggtree,
               ggstance,
               MuMIn,  # multi-model inference
               patchwork,   # putting ggplots together - you need to install via devtool
               here,    # making reading files easy
               MASS,
               cowplot #multipanel plots
)

```

### Load data: nmc_tracking_data

```{r load data}
#this is from nmc_modelling.R script
data <- read.csv(here("data", "nmc_tracking_data.csv"))
data <- data[data$outlier !=1,]

#Get herd IDs
pink <- unique(data[data$Pop_Unit == "Pink Mountain","WLH_ID"])
muskwa <- unique(data[data$Pop_Unit == "Muskwa","WLH_ID"])

```

### Function for detrending the speeds

```{r detrending the speeds}
detrend.speeds <- function(DATA, CTMM) {
  
  # estimate speeds conditional on the data
  EST <- speeds(DATA, CTMM, fast=TRUE, level=NULL)
  # null estimates of speed (no data, only model)
  EST.NULL <- speeds(CTMM, t=DATA$t, fast=TRUE, level=NULL)
  
  # reduce DOF by that of null distribution
  DOF <- EST$DOF - EST.NULL$DOF
  all(DOF>0) # check for validity... always has held so far
  DOF <- pmax(0,DOF) # haven't needed this, but just in case
  
  S2 <- EST$speed^2 + EST$VAR
  
  AVE <- (EST$DOF*S2)/DOF
  
  # Calculate CIs
  CI <- sapply(1:length(DATA$t), function(i){ctmm:::chisq.ci(AVE[i], DOF=DOF[i], level=0.95, robust=TRUE)})
  CI <- sqrt(CI)
  
  SPEEDS <- as.data.frame(t(CI))
  SPEEDS$time <- DATA$timestamp
  
  return(SPEEDS)
}

```

## Movement modelling

Key things to check for/keep in mind: 
1. movebank format is usually okay - you need to have the column individual.local.identifier, this is grabbed first
2. converted Acquisition_Date to timestamp - the timestamp column is grabbed first
3. outliers are 1's, need to be dealt with before proceeding as they mess with the model output
4. there will be warnings and/or errors - those are sanity checks
5. mark.rm = TRUE will remove the outliers

```{r movement modelling}
#Convert to telemetry object
DATA <- as.telemetry(data)

#summary(DATA) gives the sampling interval and coordinates if you want to plot the summary data
#plot(DATA) is a nice plot function to visualize the data 
#plot(DATA, col=rainbow(length(DATA)))
#class(DATA[[1]])
#test<-DATA[[1]]
#head(test)
#HDOP is the measurement error (horizontal dilution of precision)
#test@info #this gives you the identity if you need the know the individual dataset
#test@UERE #each collar manufacturer will give a different value for this, it converts the DOP value into an error in meters (DOP of 1 is 10 meters)
#another example for errors: test<-oulie(turtle[[3]]) #you will see drifting in the movement, this is a vague spot where you have to make subjective decisions, cross that bridge when you get to it

#test<-oulie(DATA[[4]]) #the farther an animal gets away from it's core use area is more red and larger points (red is less informative - those are distance based outliers)
#head(test)
#plot(test) #the blue lines connect the locations, they get thicker and more blue as the speed increases (you want to see connections between paths, but watch for the angles, they shouldn't be too tight) (blue are more informative - these are the speed based outliers)
#plot(test, units=FALSE) #changes the units

#fit error parameters to calibration data
UERE<-uere.fit(DATA[[5]][2134:3099,])
#do not run uere.fit on data where they are actually moving

#estimated error model parameters
summary(UERE)

#Drop mortality event
DATA[[5]] <- DATA[[5]][1:2134,] #[2134:3099,] #when there are lots of animals, consider just dropping it 

#uere.fit(DATA[[5]][2134:3099,])

FITS <- list()
```
:::

## R Session Information

```{r}
# pander for making it look nicer
sessionInfo() %>% pander()
```
